Minimal trap-spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-05-04 cmsb22.tex
Complete the proof of Theorem 1, update the definition of a mirror

  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) if and only if
  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S.\] If \(D_m = \emptyset\), \(S\) is conventionally \(\emptyset\).

@@

  First, we show that if \(m\) is a trap space of \(\mathcal{M}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*).

  Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{M}\) (**).

>>>>

  Let \(m\) be a subspace of Boolean model \(\mathcal{M}\). A \emph{mirror} of $m$ is a set of places $S$ in the Petri net encoding \(\mathcal{P}\) of \(\mathcal{M}\) such that:
  \[\forall v \in D_m, m(v) = 0 \Leftrightarrow p_v \in S, m(v) = 1 \Leftrightarrow \overline{p}_v \in S\] and \[\forall v \not \in D_m, p_v \not \in S, \overline{p}_v \not \in S.\] If \(D_m = \emptyset\), \(S\) is conventionally \(\emptyset\).

@@

  First, we show that if \(m\) is a trap space of \(\mathcal{M}\), then \(S\) is a conflict-free siphon of \(\mathcal{P}\) (*). If \(D_m = \emptyset\), then \(S = \emptyset\) is trivially a conflict-free siphon of \(\mathcal{P}\). Thus, we consider the case that \(D_m \neq \emptyset\) (resp. \(S \neq \emptyset\)). Assume that \(S\) is not a siphon of \(\mathcal{P}\). Then, there is a transition \(t \in T\) such that \(S\cap succ(t)\not =\emptyset\) but \(S\cap pred(t)\not =\emptyset\). In other words, there is a place \(p \in S\) such that \(p \in succ(t)\) but \(p \not \in pred(t)\). Let \(v\) be the corresponding node in \(\mathcal{M}\) of \(p\). By the characterization of the encoding~\cite{chatain2014characterization}, there is a directional arc from \(t\) to \(p\) and a directional arc from the complementary place of \(p\) to \(t\). Without loss of generality, we assume that \(p = p_v\), then there is a directional arc from \(t\) to \(p_v\) and a directional arc from \(\overline{p}_v\) to \(t\). In addition, there is also no arc or a bi-directional arc between \(t\) and another place rather than \(p_v\) and \(\overline{p}_v\). Thus, there is no connecting arc between \(t\) and any place in \(S \setminus \{p_v\}\) because \(S\cap pred(t)\not =\emptyset\). In \(\mathcal{S}_{\mathcal{M}}[m]\), a node in \(V \setminus D_m\) can receive any Boolean value. Hence, there is a state \(s \in \mathcal{S}_{\mathcal{M}}[m]\) such that \(m_s(p') = 1, \forall p' \in pred(t) \setminus \{\overline{p_v}\}\) where \(m_s\) is the corresponding marking in \(\mathcal{P}\) of \(s\). We also have \(m_s(p_v) = 0\), leading to \(m_s(\overline{p_v}) = 1\) by the characterization of the encoding~\cite{chatain2014characterization}. Now, \(t\) is enabled at marking \(m_s\). Its firing leads to a new marking \(m'_s\) such that \(m'_s(p_v) = 1\) and \(m'_s(\overline{p}_v) = 0\). Let \(s'\) be the corresponding state in \(\mathcal{M}\) of \(m'_s\). Since \(m\) is a trap space of \(\mathcal{M}\), \(s' \in \mathcal{S}_{\mathcal{M}}[m]\). Then, \(s'(v) = m(v)\), leading to \(m'_s(p_v) = 0\), which is a contradiction. Hence, \(S\) is a siphon of \(\mathcal{P}\). By the definition of a mirror, \(S\) is also a conflict-free one.

  Second, we show that if \(S\) is a conflict-free siphon of \(\mathcal{P}\), then \(m\) is a trap space of \(\mathcal{M}\) (**). By the definition of a mirror, \(m\) is a subspace of \(\mathcal{M}\). Let \(s\) be an arbitrary state in \(\mathcal{S}_{\mathcal{M}}[m]\) and \(m_s\) be its corresponding marking in \(\mathcal{P}\). By the characterization of the encoding~\cite{chatain2014characterization}, \(m_s(p) = 0, \forall p \in S\). In any marking \(m'_s\) reachable from \(m_s\) regardless of the update semantics of \(\mathcal{P}\), we have \(m'_s(p) = 0, \forall p \in S\) by the characterization of a siphon~\cite{DBLP:journals/isci/LiuB16}. Equivalently, in any state \(s'\) reachable from \(s\) regardless of the update semantics of \(\mathcal{M}\), we have \(s'(v) = s(v) = m(v), \forall v \in D_m\). Then, \(s' \in \mathcal{S}_{\mathcal{M}}[m]\). By the definition of a trap space and the arbitrariness of \(s\), \(m\) is a trap space of \(\mathcal{M}\).
